---
title: "Dataset_Comparison_CCLE_NCI"
date: "`r Sys.Date()`"
output:
  rmdformats::html_clean:
    highlight: kate

---


```{r setup, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75,
              root.dir = "./../")
```
# Overall Matrix agreement subsetted for cell lines ang genes in common and order. 
Using https://rdrr.io/cran/MatrixCorrelation/ different methods available
CCLE vs NCI = which omics agrees most between datasets = Visualisation?
Transcriptomic vs Transcriptomic
Proteome vs Proteome
Metabolome vs Metabolome

Gene/Metabolite specific comparison, is this related to abundance? Enrichment when subsetting for quartiles of abundance.
Plot against null permutation. Enrichment for essential genes? - plot Overlapping distributions

Cell Line comparison. Do some Tissues/ Cell lines agree more between them than others?

Plot in double matrix of transcripromis vs proteomics correlations ordered by pathway?


```{r Data_load_Subset}
load()

```
# Other Title