---
title: "Dataset_Comparison_CCLE_NCI"
date: "`r Sys.Date()`"
output:
  rmdformats::html_clean:
    highlight: kate
    code_folding: hide

---

# Setup

```{r setup, echo=FALSE, cache=FALSE}
pacman::p_load(knitr,rmdformats,tidyverse, MatrixCorrelation)
library(ComplexHeatmap)
library(circlize)

## Global options
options(max.print="75")
opts_chunk$set(cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75,
              root.dir = "./../")
```

# Overall Matrix agreement subsetted for cell lines ang genes in common and order. 
RPPA?
Plot in double matrix of transcripromis vs proteomics correlations ordered by pathway?


```{r Data_load_Subset}
# load(file = "./../Project_Datasets/CCLE_OMICS.rda")
# load(file = "./../Project_Datasets/NCI60_OMICS.rda")
# load(file = "./../Project_Datasets/Metabolic_Pathway_annotations.RData")
# 
# HUMAN_9606_idmapping <- read_tsv("./../Project_Datasets/HUMAN_9606_idmapping.dat",
#                                  col_names = FALSE) %>%
#     setNames(c("Uniprot", "Type", "ID"))

load(file = "./Project_Datasets/CCLE_OMICS.rda")
load(file = "./Project_Datasets/NCI60_OMICS.rda")
load(file = "./Project_Datasets/Metabolic_Pathway_annotations.RData")

HUMAN_9606_idmapping <- read_tsv("./Project_Datasets/HUMAN_9606_idmapping.dat",
                                 col_names = FALSE) %>%
    setNames(c("Uniprot", "Type", "ID"))

HUMAN_9606_idmapping_hsa <- HUMAN_9606_idmapping[str_detect(HUMAN_9606_idmapping$ID, "hsa:"),-2] %>% 
    left_join(HUMAN_9606_idmapping[HUMAN_9606_idmapping$Type == "Gene_Name", -2], by = "Uniprot") %>% 
    na.omit() %>% .[,-1] %>% setNames(c("KEGG_ID", "Gene.names"))

Correlation_pairwise <- function(x,y){
    cor(x,y,use = "pairwise", method = "pearson")
}

# Get lower triangle of the correlation matrix
  get_lower_tri<-function(cormat){
    cormat[lower.tri(cormat)] <- NA
    return(cormat)
  }
# Get upper triangle of the correlation matrix
  get_upper_tri<-function(cormat){
    cormat[upper.tri(cormat)] <- NA
    return(cormat)
  }
  
#compare 2 datasets with common Feature (rows) and Sample names (Columns)
### Change rerun value
Dataset_correlation <- function(Dataset_1,Dataset_2){
  # type = c("Transcriptomics",
  #                                                             "Proteomic",
  #                                                             "Metabolomics")
  #  type <- "Transcriptomics" 
  # Type <- case_when(
  #     type == "Transcriptomics" ~ "Gene_name",
  #     type == "Proteomic" ~ "Uniprot",
  #     type == "Metabolomics" ~ "Metabolites",
  #     TRUE ~ as.character(type))

    #Dataset_1 = CCLE_proteins
    #Dataset_2 = NCI_60_proteins
    Comparison_type <- paste(deparse(substitute(Dataset_1)),deparse(substitute(Dataset_2)))
    Dataset1 <- as.data.frame(Dataset_1)
    Dataset2 <- as.data.frame(Dataset_2)
    Common_Cell_lines <- intersect(colnames(Dataset1),colnames(Dataset2)) #Common Samples
    Common_proteins <- intersect(rownames(Dataset1),rownames(Dataset2)) #Common Features
    
        #subsetting and ordering the Datasets
    Dataset1_sub <- Dataset1 %>% .[rownames(.) %in% Common_proteins, colnames(.) %in% Common_Cell_lines]%>% 
    .[match(Common_proteins,rownames(.)),match(Common_Cell_lines,colnames(.))]
    Dataset2_sub <- Dataset2 %>% .[rownames(.) %in% Common_proteins, colnames(.) %in% Common_Cell_lines]%>% 
    .[match(Common_proteins,rownames(.)),match(Common_Cell_lines,colnames(.))] 
    Master_pathways <- Master_pathways[!duplicated(Master_pathways$ID),]
    Master_pathways_ordered <- Master_pathways[Master_pathways$ID %in% Common_proteins,] %>% .[order(.$Pathway),]
    cormat1 <- round(cor(t(Dataset1_sub[rownames(Dataset1_sub) %in% Master_pathways_ordered$ID,])),2) %>% 
      .[match(Master_pathways_ordered$ID,row.names(.)),match(Master_pathways_ordered$ID,row.names(.))] %>%
      get_lower_tri() 
    
    #%>% 
        # reshape2::melt(na.rm = T) %>% 
        # left_join(Master_pathways[!duplicated(Master_pathways$Uniprot),], by = c("Var1" = "Uniprot"))
        #left_join(SMPDB_pathways[!duplicated(SMPDB_pathways$Uniprot),c("Pathway","Uniprot")], by = c("Var1" = "Uniprot"))
    cormat1[is.na(cormat1)] <- 0
    cormat2 <- round(cor(t(Dataset2_sub[rownames(Dataset2_sub) %in% Master_pathways_ordered$ID,])),2)%>% 
      .[match(Master_pathways_ordered$ID,row.names(.)),match(Master_pathways_ordered$ID,row.names(.))] %>%
      get_upper_tri() #%>%
        # reshape2::melt(na.rm = T)%>%
        # left_join(Master_pathways[!duplicated(Master_pathways$Uniprot),], by = c("Var1" = "Uniprot"))
        #left_join(SMPDB_pathways[!duplicated(SMPDB_pathways$Uniprot),c("Pathway","Uniprot")], by = c("Var1" = "Uniprot"))
    cormat2[is.na(cormat2)] <- 0
    #Combined_df <- rbind(cormat1,cormat2)  
    # #Combined_df$value[Combined_df$Var1 == Combined_df$Var2] <- NA
    # 
    # Feature_order <- Combined_df[!duplicated(Combined_df$Var1),] %>% 
    #   .[order(.$Pathway),"Var1"]
    # 
    # ### ComplexHeatMAP
    # test1 <- Master_pathways_ordered[order(Master_pathways_orderezd$Uniprot),]
    Metabolic_matrix <- as.data.frame(cormat1+cormat2) %>%       as.matrix()
      
    
    set.seed(15887) # to set random generator seed
    cl <- colors(distinct = TRUE)
  
    
    Pathway_colours <- sample(cl, length(unique(Master_pathways_ordered$Pathway))) %>% setNames(unique(Master_pathways_ordered$Pathway))
    Metabolic_matrix[Metabolic_matrix ==2] <- 1
    ha = HeatmapAnnotation(df = Master_pathways_ordered$Pathway,
                           col = list(df = Pathway_colours))
    hmap <- Heatmap(Metabolic_matrix, name = Comparison_type, 
            top_annotation = ha ,
            left_annotation = rowAnnotation(df = Master_pathways_ordered$Pathway,col = list(df = Pathway_colours)),
            cluster_rows = F, cluster_columns = F, show_column_names  = F, show_row_names = F, 
            row_title = paste0("top = ", deparse(substitute(Dataset_1))), 
            column_title = paste0("bottom = ", deparse(substitute(Dataset_2))),
            column_title_side = "bottom"
            )
    


    # ### GGPLOT
    # Combined_df %>%
    #     subset(Var1 %in% Master_pathways$Uniprot & Var2 %in% Master_pathways$Uniprot) %>% na.omit() %>%
    #     ggplot(aes(factor(Var2, levels = Feature_order), factor(Var1,levels = Feature_order), fill = value))+
    #     geom_tile(color = "white")+
    #     scale_fill_gradient2(low = "blue", high = "red", mid = "white", na.value = "black",
    #                          midpoint = 0, limit = c(-1,1), space = "Lab", 
    #                          name="Pearson\nCorrelation") +
    #     theme_minimal()+ 
    #     theme(axis.text.x = element_blank(), axis.text.y = element_blank())+
    #     coord_fixed() + labs(x = deparse(substitute(Dataset_1)),deparse(substitute(Dataset_2)))#+
    # facet_grid(Pathway~., scales = "free_x")
    
    
    #Correlating corresponding Samples Correlation of only paired values which allows for NA values
    Cell_line_correlation <- map2_dbl(
        Dataset1_sub, 
        Dataset2_sub, 
        Correlation_pairwise) 
    
    Cell_line_perm <- purrr::rerun(10, 
                                   map2_dbl(
                                       Dataset1_sub[,sample(ncol(Dataset1_sub))],#Shuffle column-wise and performing sample correlation as above
                                       Dataset2_sub,
                                       Correlation_pairwise)) %>% unlist()
    
    gene_correlation <- pmap(
        list(pmap(Dataset1_sub, c, use.names = T), # unlisting feature rows into vectors into lists and correlating
             pmap(Dataset2_sub, c, use.names = T)),
        Correlation_pairwise) %>% unlist() %>% setNames(Common_proteins) 
    
    gene_correlation_perm <- purrr::rerun(10, #permutation by row shuflfing and correlation
                                          pmap(list(pmap(Dataset1_sub[sample(nrow(Dataset1_sub)),], 
                                                         c, use.names = T), #Shuffle row-wise
                                                    pmap(Dataset2_sub, c, use.names = T)),
                                               Correlation_pairwise) %>% unlist()) %>% unlist()
    
    # Imputing both datasets with Media to derive overall correlation
    Dataset1_sub_imp <- naniar::impute_median_all(Dataset1_sub)
    Dataset2_sub_imp <- naniar::impute_median_all(Dataset2_sub)
    
    Matri_Corr <- MatrixCorrelation::r1(Dataset1_sub_imp,Dataset2_sub_imp) # whole matrix correlation
    
    Matrix_perm <- purrr::rerun(10, #whole matrix permutation of Dataset 1 col-wise shuffle and Dataset 2 row-wise shuffle and correlation
                            MatrixCorrelation::r1(Dataset1_sub_imp[,sample(ncol(Dataset1_sub_imp))],
                                                  Dataset2_sub_imp[sample(nrow(Dataset2_sub_imp)),])) %>% 
        unlist() # permutation
    
    Hist_plot <- function(Data,Perm){
        Sample_Corr <- data.frame(Samples = c(Data,unlist(Perm)),
                              Type = rep(c("Data","Perm"), times = c(length(Data),length(unlist(Perm)))))
        p <- ggplot2::ggplot(Sample_Corr,aes(Samples,after_stat(density), colour = Type)) + geom_freqpoly(binwidth = 0.01) +
            ggtitle(paste(Comparison_type,deparse(substitute(Data))))
    }
    
    Feature_quartiles <- rowMeans(Dataset2_sub, na.rm = T) %>% data.frame(Feature = names(.),
                                                           Mean_Abundance = .) %>%
        mutate(quartile = as.factor(ntile(Mean_Abundance, 4))) %>% left_join(stack(gene_correlation), by = c("Feature" = "ind"))
    
    Feature_expression_cor_plot <- ggplot(Feature_quartiles, aes(values,after_stat(density), colour =quartile)) +
        geom_freqpoly(binwidth = 0.1) +
        ggtitle(paste("Feature Expression",Comparison_type))
    
    Sample_lineage <-  Cell_line_correlation %>% data.frame(Sample = names(.), Correlation = .) %>%
      left_join(sample_info[,c("stripped_cell_line_name", "lineage")], by = c("Sample" = "stripped_cell_line_name")) %>%
      left_join(Derms, by = c("lineage" = "Cancer_Lineage" ))
    
    Sample_lineage_cor_plot <- ggplot(Sample_lineage, aes(Correlation,after_stat(density), colour = Derm)) +
        geom_freqpoly(binwidth = 0.01) +
        ggtitle(paste("Sample Lineage",Comparison_type))
    
    
    list(Sample_Corr = list(Sample_lineage, Cell_line_perm, 
                            Hist_plot(Cell_line_correlation,Cell_line_perm),
                            Sample_lineage_cor_plot),
         Feature_Corr = list(Feature_quartiles, gene_correlation_perm, 
                             Hist_plot(gene_correlation, gene_correlation_perm), Feature_expression_cor_plot),
         Matrix_Corr= list(Matri_Corr, Matrix_perm, 
                           Hist_plot(Matri_Corr,Matrix_perm)),
         Heatmap= hmap)
    
}
# <- openxlsx::read.xlsx("./../Project_Datasets/Lineage_tissue_types.xlsx", cols = c(1,6)) %>% na.omit()
Derms <- openxlsx::read.xlsx("./Project_Datasets/Lineage_tissue_types.xlsx", cols = c(1,6)) %>% na.omit()
#Plotting_Essentiality


### Transcriptomic Comparison ###
CCLE_NCI_transcr <- Dataset_correlation(RNA_seq,NCI_60_RNA)
Essential_vs_Not_Genes <- CCLE_NCI_transcr$Feature_Corr[[1]] %>%
    mutate(Essentiality = if_else(Feature %in% Essential_genes, "Essential", "Not_Essential"))

ggplot(Essential_vs_Not_Genes, aes(values,after_stat(density), colour = Essentiality)) +
        geom_freqpoly(binwidth = 0.1) +
        ggtitle(paste("mRNA essentiality Correlation"))
### Proteomic Comparison ###
CCLE_NCI_prot <- Dataset_correlation(CCLE_proteins,NCI_60_proteins)



### Metabolic Comparison ###
#Metabolite_mapping <- read.csv("./../Project_Datasets/CCLE_metabolites_mapped.csv")

Metabolite_mapping <- read.csv("./Project_Datasets/CCLE_metabolites_mapped.csv")
rownames(NCI_60_metabolites) <- rownames(NCI_60_metabolites) %>% str_replace_all("HMDB", "HMDB00")

Common_Metabolites <- Metabolite_mapping[(Metabolite_mapping$KEGG %in% rownames(NCI_60_metabolites)) |
                                          (Metabolite_mapping$HMDB %in% rownames(NCI_60_metabolites)) ,] %>%
  .[!(.$Query %in% c("NAD","NADP")),]


NCI_60_metabolites_common <- NCI_60_metabolites[rownames(NCI_60_metabolites) %in% c(Common_Metabolites$HMDB,Common_Metabolites$KEGG),]
tt1=rownames(NCI_60_metabolites_common)[str_detect(rownames(NCI_60_metabolites_common),"^H")]
rownames(NCI_60_metabolites_common)[str_detect(rownames(NCI_60_metabolites_common),"^H")] <- Metabolite_mapping[Metabolite_mapping$HMDB %in% tt1,] %>%
  .[match(tt1, .$HMDB),] %>% .$KEGG

CCLE_Met_Common <- Metabolites[rownames(Metabolites) %in% Common_Metabolites$Query,]


rownames(CCLE_Met_Common) <- Common_Metabolites$KEGG[match(rownames(CCLE_Met_Common),Common_Metabolites$Query)]

CCLE_NCI_met <- Dataset_correlation(CCLE_Met_Common,NCI_60_metabolites_common)

omics <- list(Transcriptome = CCLE_NCI_transcr,
              Proteome = CCLE_NCI_prot,
              Metabolome = CCLE_NCI_met)

```

# Sample correlation

```{r Sample_cor}
map(omics, ~.x$Sample_Corr[[3]])

```

# Feature correlation

```{r Feature_Corr}
map(omics, ~.x$Feature_Corr[[3]])

```

# Matrix correlation

```{r Matrix_Corr}
map(omics, ~.x$Matrix_Corr[[3]])
```

# Sample Lineage

```{r Sample_lineage}
map(omics, ~.x$Sample_Corr[[4]])

```


# Feature Gene expression Corr

```{r Feature_expre}
map(omics, ~.x$Feature_Corr[[4]])
```

1 is the bottom quartile and 4 the top quartile


# Heatmaps

```{r heatmaps}
map(omics, ~.x$Heatmap)
```
Here we see that the proteins might actually be more biologically relevant than the mRNA and the 2 datasets are comparable. The CCLE proteome has more NAs but this is not shown here due to 'pairwise' correlation. But when the data is present, the comparison is good

