---
title: "Drug Map MOFA w Blood & Skin"
date: "`r Sys.Date()`"
output:
    html_document:
      code_folding: hide
      toc: true
      toc_depth: 2
      highlight: tango


---

# Setup

```{r setup,  echo=FALSE, cache=FALSE, warning= FALSE, message=FALSE}
pacman::p_load(knitr,tidyverse, here)
#devtools::install_github("bioFAM/MOFA2", build_opts = c("--no-resave-data --no-build-vignettes"))

## Global options
options(max.print="75")
opts_chunk$set(cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
#opts_knit$set(root.dir = "./../")
Files_to_load <- c(here::here("Project_Datasets","Metabolic_Pathway_annotations.RData"),
                   here::here("Project_Datasets","Drug_sensitivity.RData"))
                  
HUMAN_9606_idmapping_file <- here::here("Project_Datasets","HUMAN_9606_idmapping.dat")

if(intersect(map(Files_to_load, cgwtools::lsdata) %>% unlist(),
             ls()) %>% length() == 0){
  purrr::walk(Files_to_load, load, env =.GlobalEnv)
} else {
  stop("Variable names in environment which clash with RData loading Names")
             }

#Mapping of Uniprots to all others sort of identifiers
HUMAN_9606_idmapping <- read_tsv(HUMAN_9606_idmapping_file,
                                 col_names = FALSE) %>%
    setNames(c("Uniprot", "Type", "ID"))

```
ABCB1 belongs to Transport of small molecules and is anticorrelated with tepoxalin[K89714990] but the drug prediction accuracy is only 0.552402 with a large SD (this pathway has 685 features which might be diluting the signal)
EGFR is a direct target and predictor of gefitinib[K64052750]. So when checking for "Constitutive Signaling by Ligand-Responsive EGFR Cancer Variants", small pathway containing EGFR, then the predictions for ERBB2 pathways become very sharp ~ 4-5 other drugs also target the same pathways and are predicted well = 	
PD-168393 an irreversible inactivator of EGFR and ERBB2
AEE788 is a multitargeted human epidermal receptor (HER)
poziotinib is irreversible tyrosine kinase inhibitor targeting EGFR and HER2
BMS-599626 is a highly selective and potent inhibitor of HER1 and HER2
canertinib it is an irreversible tyrosine-kinase inhibitor with activity against EGFR
Afatinib It is mainly used to treat cases of NSCLC that harbour mutations in the epidermal growth factor receptor (EGFR) gene

```{r Drug_predictions}
Importing_pathways <- function(x){
     # x = list.files(path = here::here("Project_Datasets",
     #                          "Pathway_analysis_cluster"),
     #                          pattern = "*wo_blood_skin.csv")[1]
  pathway <- read_csv(here::here("Project_Datasets",
                             "Pathway_analysis_cluster",x),
                      col_names = F) %>%
    dplyr::select(-X1,-X8)

  if(ncol(pathway) == 7){
    return(pathway)
  }else{
    stop("Irregular columns")
  }
}
Importing_pathways_safe <- safely(Importing_pathways)
library(future)
plan(multisession, workers = 8)
Cluster_pathways_f <-  future({purrr::map(list.files(path = here::here("Project_Datasets",
                             "Pathway_analysis_cluster"),
           pattern = ".csv") %>% .[!str_detect(.,"wo_blood_skin")],Importing_pathways_safe)},
           seed = T)

Cluster_pathways <- value(Cluster_pathways_f) %>%
  map("result") %>%  compact()%>% 
  purrr::reduce(rbind)

plan(sequential)
Cluster_pathways <- Cluster_pathways%>%
  setNames(c( "Accuracy", "Kappa", "AccuracySD", "KappaSD","Drug","model","Pathway"))

Best_prediction <- Cluster_pathways %>%
  group_by(Drug) %>% summarise(max_accuracy = max(Accuracy))
Best_prediction %>% pull(max_accuracy) %>% hist(main="Histogram for Best prediction per drug")
Unpredicted_Drugs <-  Best_prediction %>%
  filter(max_accuracy<0.75) %>% pull(Drug)

Cluster_pathways2 <- Cluster_pathways %>% 
  filter(!(Drug %in% Unpredicted_Drugs))  %>%
  inner_join(drug_targets[,1:2], by = "Drug")

Cluster_pathways3 <- Cluster_pathways2 %>% 
  group_by(Pathway,Reactome_Pathway) %>% 
  add_tally(name = "Drugs_per_pathway")%>% 
  filter(Drugs_per_pathway>3) %>% ungroup() %>%
  group_split(Pathway) %>%
  setNames(.,map_chr(.,~(unique(.x[['Pathway']]))))


Drug_prediction_plot <- function(x){
  #x = Cluster_pathways2 %>% filter(Pathway == "Constitutive Signaling by Ligand-Responsive EGFR Cancer Variants")
  #x
  high_Pathways <- x %>% group_by(Reactome_Pathway,Pathway) %>%
    summarise(mean_drug = mean(Accuracy)) %>%  ungroup %>%
    filter(mean_drug > (quantile(.$mean_drug,prob = seq(0, 1, length = 50), 
                                 type = 5, na.rm =T)[[49]])) %>% 
    pull(Reactome_Pathway)
  Pathway_of_interest = "Activation of BAD and translocation to mitochondria"
  test <- x %>% filter(Reactome_Pathway %in% high_Pathways) 
  if(nrow(test) <2){
    unique(x$Pathway)
  }else{
  test %>%
    ggplot(aes(x = Accuracy, fill = Reactome_Pathway, colour = Reactome_Pathway))+ ggtitle(unique(x$Pathway)) +
  #geom_histogram(alpha=0.6, position = 'identity') #theme(legend.position = "none") 
  geom_density(alpha = 0.05)+theme_bw() #+ theme(legend.position = "none") 
    }
}
map(Cluster_pathways3[60:80],Drug_prediction_plot)
chosen_pathways <- c("Constitutive Signaling by Ligand-Responsive EGFR Cancer Variants",
                     "Amplification of signal from the kinetochores",
                     "Autophagy",
                     "Asymmetric localization of PCP proteins")
map(Cluster_pathways3[chosen_pathways],Drug_prediction_plot)
```

